---
title: 'Assignment 1: Grandparents on duty'
author: "by Cristina Davila Gonzalez"
date: "9/4/2020"
output: html_document

---

*On this year's Back to School season, most of public schools worldwide have decided to go remote. Coming from a current context of diaspora and mass migration of young/adult professionals,  where sometimes elders are left in charge of their grandchildren, I decided to take a look at similar family scenarios in the U.S. where grandparents are a key factor in making remote-learning feasible for their underage relatives.*

```{r, message = FALSE}
library(tidycensus)
library(tidyverse)
```

# Variables Used

For this exercise I used multi-level variables from the **American Community Survey 2018**, adding person-level and household-level data (both from the ACS 1-year data) to eventually generate supporting information on the research question, specifically in the state of **Illinois**.

##   Person-Level Data Analysis 

1. **Age** (AGEP) (continuous)
2. **School Enrollment** (SCH) (categorical)
3. **Grade Level Attending** (SCHG) (categorical)
4. **Self-care difficulty** (FDDRSP) (categorical)

As I developed my dataset, I decided to apply a filter to narrow my search to **underage enrolled students** (<18 y/o)

```{r load person-level data analysis, message=FALSE, results='hide'}
person_data <- get_pums(variables = c("AGEP", 
                            "SCH", 
                            "SCHG", 
                            "FDDRSP"),
                        state = "IL", 
                        year = 2018, 
                        survey = "acs1",
                        recode = TRUE) %>%
  filter(AGEP < 18) %>%
  select(AGEP, SCH_label, SCHG_label, FDDRSP_label)

```

```{r}
person_data
```
My person-level dataset has 25,993 rows, which mean my analysis will include 25,993 students.

##   Household-Level Data Analysis 

6. **Presence of persons under 18** (R18) (categorical)
7. **Access to internet connection** (ACCESS) (categorical)
8. **Grandparent-headed household with no parent present** (NPP) (categorical)
9. **Limited English-speaking household** (LNGI) (categorical)

```{r load household-level data, message=FALSE, results = 'hide',}
hh_data <- get_pums(variables = c("R18",
                        "ACCESS",
                        "NPP", 
                        "LNGI"),
                    
                        state = "IL",
                        year = 2018,
                        survey = "acs1",
                        recode = TRUE) %>%
  
  select(R18_label, ACCESS_label, NPP_label, LNGI_label)
```

```{r}
hh_data
```

My household-level dataset has 126,456 rows, which mean my analysis will include 126,456 households.

For both datasets, I'll save them as a CSV files so I can use them later.

```{r}
write_csv(person_data, "u18people.csv")
write_csv(hh_data, "households.csv")
```
