---
title: 'Assignment 1: Grandparents on duty'
author: "by Cristina Davila Gonzalez"
date: "9/4/2020"
output: html_document

---

*On this year's Back to School season, most of public schools worldwide have decided to go remote. Coming from a current context of diaspora and mass migration of young/adult professionals,  where sometimes elders are left in charge of their grandchildren, I decided to take a look at similar family scenarios in the U.S. where grandparents are a key factor in making remote-learning feasible for their underage relatives.*

```{r, message = FALSE}
library(tidycensus)
library(tidyverse)
library(dplyr)
```

# Variables Used

For this exercise I used person-level variables from the **American Community Survey 2018** (ACS 1-year data) to eventually generate supporting information on the research question, specifically in the state of **Illinois**. As I developed my analysis on the selected variables, I decided to generate two datasets; one with **underage enrolled students** and another one considering **grandparents of 50 years old and up** in charge of their grandchildren.

1. **Age** (AGEP) (continuous)
2. **School Enrollment** (SCH) (categorical)
3. **Grade Level Attending** (SCHG) (categorical)
4. **Self-care difficulty** (FDDRSP) (categorical)
5. **Grandparents responsible for grandchildren** (GCR) (categorical)
6. **Total person's income** (PINCP) (continuous)

For this first dataset, I'm only considering variables *1,2,3 and 4*, applying a filter to the *AGEP* variable to narrow my search to individuals below the age of 18.

`filter(AGEP < 18) %>%`


```{r load person-level data, message=FALSE, results='hide'}
personU18_data <- get_pums(variables = c("AGEP", 
                            "SCH", 
                            "SCHG", 
                            "FDDRSP"),
                        state = "IL", 
                        year = 2018, 
                        survey = "acs1",
                        recode = TRUE) %>%
  filter(AGEP < 18) %>%
  select(AGEP, SCH_label, SCHG_label, FDDRSP_label)

```

```{r}
personU18_data
```
My first dataset has 25,993 rows, which mean my analysis will include **25,993 underage individuals**.


For my second dataset, I'm considering variables *1,4,5 and 6*, applying a filter to the **AGEP** variable to narrow my search to individuals of age 50 and up, as it is the average age at first-grandchild in the U.S., according to the **American Association of Retired Persons 2018 Survey**. 
https://www.aarp.org/content/dam/aarp/research/surveys_statistics/life-leisure/2019/aarp-grandparenting-study.doi.10.26419-2Fres.00289.001.pdf

`filter(AGEP > 64) %>%`


I'm also setting up my dataset so it discards all N/A results in my **GCR** variable (*5*), with a N/A troubleshoot resource shared to me by Adriana Lasso-Harrier, also supported by our TF, Sara Tufter.

`O50_wgrandchild <- personO50_data[personO50_data$GCR_label!= "N/A (less than 30 years/institutional GQ/grandparent not living with grandchild)",]`


```{r message=FALSE, include=TRUE, results='hide'}
personO50_data <- get_pums(variables = c("AGEP", 
                            "GCR",
                            "PINCP",
                            "FDDRSP"),
                        state = "IL", 
                        year = 2018, 
                        survey = "acs1",
                        recode = TRUE) %>%
  filter(AGEP > 49) %>%
  select(AGEP, GCR_label,PINCP, FDDRSP_label)

O50_wgrandchild <- personO50_data[personO50_data$GCR_label!= "N/A (less than 30 years/institutional GQ/grandparent not living with grandchild)",]
```


```{r}
O50_wgrandchild
```

My second dataset has 2001 rows, which mean my analysis will include **2001 grandparents of age 50 and above**.

For both datasets, I'll save them as a CSV files so I can use them later.

```{r}
write_csv(personU18_data, "u18people.csv")
write_csv(O50_wgrandchild, "O65people.csv")
```
