---
title: "Assignment 2: Calculations"
author: "Cristina Davila Gonzalez"
date: "9/14/2020"
output: html_document

---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
```

# Assignment 1 datasets
```{r message=FALSE, results='hide'}
u18people_data <-read_csv("u18people.csv")
o50people_data <-read_csv("O50people.csv")
```

even though this is a calculations-only document, I've decided to properly identify all of them for future consultations.

# Continuous variables - Summary & Individual Parts

## 1. **Age**

    1.1 Age of population under 18 within tract: AGEP in u18people_data

```{r message=FALSE, results='hide'}
u18people_summary <- summary(u18people_data$AGEP)
u18people_summary
```

Mean
```{r message=FALSE, results='hide'}
mean(u18people_data$AGEP)
```

Sample Standard Deviation 
```{r message=FALSE, results='hide'}
sd(u18people_data$AGEP)
```

95-percent Confidence Interval
```{r message=FALSE, results='hide'}
conf_int <- t.test(u18people_data$AGEP)
conf_int
```

Center of 95-percent Confidence Interval
```{r message=FALSE, results='hide'}
conf_int$estimate
```

Histogram
```{r message=FALSE, results='hide'}
ggplot(u18people_data, aes(x = AGEP)) +
  geom_histogram() +
 scale_x_continuous(name = "Age of people under 18 within tract") +
   theme_minimal()
```

    1.2 **Age** of population over 50 within tract:  AGEP in o50people_data
    
```{r message=FALSE, results='hide'}

o50people_summary <- summary(o50people_data$AGEP)
o50people_summary
```

Mean
```{r message=FALSE, results='hide'}
mean(o50people_data$AGEP)
```

Sample Standard Deviation 
```{r message=FALSE, results='hide'}
sd(o50people_data$AGEP)
```

95-percent Confidence Interval
```{r message=FALSE, results='hide'}
conf_int <- t.test(o50people_data$AGEP)
conf_int
```

Center of 95-percent Confidence Interval
```{r message=FALSE, results='hide'}
conf_int$estimate
```

Histogram
```{r message=FALSE, results='hide'}
ggplot(o50people_data, aes(x = AGEP)) +
  geom_histogram() +
 scale_x_continuous(name = "Age of people over 50 within tract") +
   theme_minimal()
```

## 6. **Total person's income**: PINCP in o50people_data

```{r message=FALSE, results='hide'}

o50people_summary <- summary(o50people_data$PINCP)
o50people_summary
```

Mean
```{r message=FALSE, results='hide'}
mean(o50people_data$PINCP)
```

Sample Standard Deviation 
```{r message=FALSE, results='hide'}
sd(o50people_data$PINCP)
```

95-percent Confidence Interval
```{r message=FALSE, results='hide'}
conf_int <- t.test(o50people_data$PINCP)
conf_int
```

Center of 95-percent Confidence Interval
```{r message=FALSE, results='hide'}
conf_int$estimate
```

Histogram
```{r message=FALSE, results='hide'}
ggplot(o50people_data, aes(x = PINCP)) +
  geom_histogram() +
 scale_x_continuous(name = "Total person's income") +
   theme_minimal()
```

# Categorical variables - Proportions

## 2. **School Enrollment**: SCH_label in u18people_data

Possible Values
```{r message=FALSE, results='hide'}
unique(u18people_data$SCH_label)
```

Number of Observations
```{r message=FALSE, results='hide'}
table(u18people_data$SCH_label)
```


Proportions of Observations
```{r message=FALSE, results='hide'}
table(u18people_data$SCH_label) / sum(table(u18people_data$SCH_label))
```

95-percent Confidence Interval of each category

 N/A (less than 3 years old)
```{r message=FALSE, results='hide'}
t.test(u18people_data$SCH_label == "N/A (less than 3 years old)")
```

 No, has not attended in the last 3 months
```{r message=FALSE, results='hide'}
t.test(u18people_data$SCH_label == "No, has not attended in the last 3 months)")
```
 
 Yes, private school or college or home school
```{r message=FALSE, results='hide'}
t.test(u18people_data$SCH_label == "Yes, private school or college or home school")
```

 Yes, public school or public college 
```{r message=FALSE, results='hide'}
t.test(u18people_data$SCH_label == "Yes, public school or public college")
```

## 3. **Grade Level Attending**: SCHG_label in u18people_data

Possible Values
```{r message=FALSE, results='hide'}
unique(u18people_data$SCHG_label)
```

Number of Observations
```{r message=FALSE, results='hide'}
table(u18people_data$SCHG_label)
```

Proportions of Observations
```{r message=FALSE, results='hide'}
table(u18people_data$SCHG_label) / sum(table(u18people_data$SCHG_label))
```

95-percent Confidence Interval of each category

 N/A (not attending school)
```{r message=FALSE, results='hide'}
t.test(u18people_data$SCHG_label == "N/A (not attending school)")
```

 Nursery school/preschool
```{r message=FALSE, results='hide'}
t.test(u18people_data$SCHG_label == "Nursery school/preschool")
```
 
 Kindergarten
```{r message=FALSE, results='hide'}
t.test(u18people_data$SCHG_label == "Kindergarten")
```

 Grade 1 
```{r message=FALSE, results='hide'}
t.test(u18people_data$SCHG_label == "Grade 1")
```

 Grade 2 
```{r message=FALSE, results='hide'}
t.test(u18people_data$SCHG_label == "Grade 2")
```

 Grade 3 
```{r message=FALSE, results='hide'}
t.test(u18people_data$SCHG_label == "Grade 1")
```

 Grade 4 
```{r message=FALSE, results='hide'}
t.test(u18people_data$SCHG_label == "Grade 4")
```


 Grade 5 
```{r message=FALSE, results='hide'}
t.test(u18people_data$SCHG_label == "Grade 5")
```

 Grade 6 
```{r message=FALSE, results='hide'}
t.test(u18people_data$SCHG_label == "Grade 6")
```

 Grade 7 
```{r message=FALSE, results='hide'}
t.test(u18people_data$SCHG_label == "Grade 7")
```

 Grade 8 
```{r message=FALSE, results='hide'}
t.test(u18people_data$SCHG_label == "Grade 8")
```

 Grade 9 
```{r message=FALSE, results='hide'}
t.test(u18people_data$SCHG_label == "Grade 9")
```

 Grade 10 
```{r message=FALSE, results='hide'}
t.test(u18people_data$SCHG_label == "Grade 10")
```

 Grade 11 
```{r message=FALSE, results='hide'}
t.test(u18people_data$SCHG_label == "Grade 11")
```

 Grade 12 
```{r message=FALSE, results='hide'}
t.test(u18people_data$SCHG_label == "Grade 12")
```

## 4. **Self-care difficulty** 

    4.1 Self-care difficulty in population under 18 within tract: FDDRSP_label in u18people_data

Possible Values
```{r message=FALSE, results='hide'}
unique(u18people_data$FDDRSP_label)
```

Number of Observations
```{r message=FALSE, results='hide'}
table(u18people_data$FDDRSP_label)
```

Proportions of Observations
```{r message=FALSE, results='hide'}
table(u18people_data$FDDRSP_label) / sum(table(u18people_data$FDDRSP_label))
```

95-percent Confidence Interval of each category

 No
```{r message=FALSE, results='hide'}
t.test(u18people_data$FDDRSP_label == "No")
```

 Yes
```{r message=FALSE, results='hide'}
t.test(u18people_data$FDDRSP_label == "Yes")
```

    4.2 Self-care difficulty in population over 50 within tract: FDDRSP_label in o65people_data
    
Possible Values
```{r message=FALSE, results='hide'}
unique(o50people_data$FDDRSP_label)
```

Number of Observations
```{r message=FALSE, results='hide'}
table(o50people_data$FDDRSP_label)
```

Proportions of Observations
```{r message=FALSE, results='hide'}
table(o50people_data$FDDRSP_label) / sum(table(o50people_data$FDDRSP_label))
```

95-percent Confidence Interval of each category

 No
```{r message=FALSE, results='hide'}
t.test(o50people_data$FDDRSP_label == "No")
```

 Yes
```{r message=FALSE, results='hide'}
t.test(o50people_data$FDDRSP_label == "Yes")
```
    
## 5. **Grandparents responsible for grandchildren**: 

Possible Values
```{r message=FALSE, results='hide'}
unique(o50people_data$GCR_label)
```

Number of Observations
```{r message=FALSE, results='hide'}
table(o50people_data$GCR_label)
```

Proportions of Observations
```{r message=FALSE, results='hide'}
table(o50people_data$GCR_label) / sum(table(o50people_data$GCR_label))
```

95-percent Confidence Interval of each category

 No
```{r message=FALSE, results='hide'}
t.test(o50people_data$GCR_label == "No")
```

 Yes
```{r message=FALSE, results='hide'}
t.test(o50people_data$GCR_label == "Yes")
```

